\section{Discussion}
\label{sec:discussion}

The central practical message of this study is that a pump-laser
datasheet splits into two distinct damage channels, quantified row by
row in Table~\ref{tab:budget}. Spectral specifications---center-frequency
stability and longitudinal-mode structure---mix inequivalent Schmidt
decompositions in the ensemble average of Eq.~(\ref{eq:rho}) and
degrade the heralded purity. Statistical specifications---intensity
noise---rescale each realization without changing its spectral shape,
leaving the purity invariant while the double-pair boost inflates
accidental coincidences through the multi-pair channel analyzed in
Ref.~\cite{takeoka2015}. Selecting a pump therefore means checking
both columns: a spectrally quiet laser with high RIN delivers pure but
noisy heralds, while an intensity-stable laser with a wandering center
frequency delivers clean statistics on a mixed state.

The budgets also show where a designer has leverage. Jitter tolerance
scales as $1/L$, a scaling we now demonstrate directly rather than
infer: across crystals of 20, 10, and 5~mm (pump re-matched in each
case) the 5\% budgets of $31.3 \pm 0.4$, $62.7 \pm 0.7$, and
$125 \pm 1$~GHz~rms stand in the ratios $1:2.00:4.00$, and
through-origin linear fits in $1/L$ describe all three drop levels
within twice the seed error (Fig.~\ref{fig:scaling}; uncertainties
throughout are mean $\pm$ SEM over 16 Monte-Carlo seeds, 8 seeds for
the RIN sweep). Halving the
crystal from 10~mm to 5~mm (pump re-matched to 0.328~THz) thus
doubles every budget, from
$27.7 \pm 0.3$/$62.7 \pm 0.7$/$91 \pm 1$~GHz~rms to
$54.1 \pm 0.9$/$125 \pm 1$/$181 \pm 3$~GHz~rms at the 1\%/5\%/10\%
levels (Fig.~\ref{fig:jitter}), at the price of a lower pair rate
that our model does not capture. The ideal purity itself is unchanged by this
trade: the first-order JSA of Eq.~(\ref{eq:jsa}) is invariant under
the rescaling $L \to L/2$, $\sigma \to 2\sigma$ up to a common
dilation of the frequency axes, which leaves the Schmidt
spectrum---and hence the purity---exactly unchanged, and the
slightly higher value quoted for the 5~mm sweep (0.818 versus 0.810)
is a numerical windowing artifact---the fixed frequency window
truncates relatively more of the sinc side lobes of the twice-wider
5~mm JSA---not a physical improvement. Bandwidth, by contrast, is the easy specification:
the 1\% window spans 0.159--0.202~THz, a full width of roughly 26\% of
the reference bandwidth, and the 10\% window spans 0.114--0.281~THz.
Center stability is the hard one: the 1\% budget of
$27.7 \pm 0.3$~GHz~rms corresponds, via Eq.~(\ref{eq:dlam}) at
$\lambda = 775$~nm, to a wavelength stability of only 55.5~pm
($\approx 0.06$~nm). A pump may exceed its nominal bandwidth by tens
of percent with negligible cost---though narrowing below the nominal
value is penalized sooner, because the reference bandwidth sits near
the lower edge of the tolerance window---but its center must hold to
a sub-0.1-nm tolerance.

\subsection*{Reading a datasheet}

The budgets above are stated in the model's units---GHz~rms of center
jitter, THz of amplitude bandwidth, percent rms of intensity
noise---whereas manufacturers print nanometer spectral widths,
temperature and current coefficients, side-mode suppression ratios,
and integrated noise percentages. The translation layer of
Sec.~\ref{sec:model}, Eqs.~(\ref{eq:fwhm})--(\ref{eq:fsr}), closes
this gap, and Table~\ref{tab:datasheet} applies it line by line to
three real products, one from each class of pump laser a designer is
likely to consider. As a worked anchor for the bandwidth line: the
reference $\sigma/2\pi = 0.164$~THz corresponds, via
Eq.~(\ref{eq:fwhm}), to an intensity FWHM of 0.273~THz, i.e.\ 0.55~nm
at 775~nm by Eq.~(\ref{eq:dlam}).

\begin{table*}
\centering
\input{tables/datasheet_map}
\caption{Reading three real pump-laser datasheets against the
imperfection budgets. Each printed specification (first two columns)
is converted into the model's units (third column) with the
translation layer of Sec.~\ref{sec:model},
Eqs.~(\ref{eq:fwhm})--(\ref{eq:fsr}), and compared against the
Monte-Carlo budgets of Table~\ref{tab:budget} (last column). Lines the
manufacturers do not publish are marked as such---for the free-running
diode the omissions are themselves the finding. Sources are the public
datasheets cited in each block header.}
\label{tab:datasheet}
\end{table*}

The free-running Fabry--Perot diode (Thorlabs L785P090) is the
cautionary case: its datasheet publishes \emph{none} of the
budget-relevant lines---no spectral width, no mode spacing, no
temperature coefficient, no RIN, no side-mode suppression ratio. The
printed 775--795~nm center wavelength is unit-to-unit selection, not
drift. A Fabry--Perot diode is multimode by construction, and
Sec.~\ref{sec:results} shows that a single additional equal-power
mode already costs 17\% of the purity, more than the entire 10\%
budget, so the class fails before any number is checked. The deeper
lesson is the absence itself: such a laser cannot even be budgeted
from its datasheet, and is unusable for this application without full
external spectral characterization.

The stabilized DFB diode (eagleyard EYP-DFB-0780) is the opposite
extreme: every spectral-stability line it publishes passes with
enormous margin. Its 1~MHz maximum linewidth enters the quasi-static
model only as a 0.5~MHz bound, more than four orders of magnitude
below the $27.7 \pm 0.3$~GHz budget (a factor of
$\sim 2\times 10^{-5}$). Its temperature coefficient,
$d\lambda/dT = 0.06$~nm/K, translates by Eq.~(\ref{eq:dlam}) to
29.9~GHz/K, so the full 1\% jitter budget---read as a uniform drift
via Eq.~(\ref{eq:pp})---equals holding the case temperature to
$\pm 1.6$~K (and, through $d\lambda/dI = 0.003$~nm/mA, the drive
current to $\pm 32$~mA), ordinary servo performance. Its side-mode
suppression of at least 30~dB is a side-line weight of $10^{-3}$, a
negligible admixture. Yet this spectrally exemplary laser is the
wrong tool for the present purpose: a CW pump with MHz linewidth sits
five orders of magnitude below the 0.159--0.202~THz bandwidth window,
and the purity collapses toward the strongly anti-correlated quasi-CW
regime---already $P = 0.04$ at the 5~GHz narrowband anchor of
Sec.~\ref{sec:results}, and falling roughly in proportion to
bandwidth below that. The DFB class
is the wrong choice for pure heralded photons, though precisely the
right one for CW-pumped entanglement sources, where center stability
is what matters.

The picosecond mode-locked oscillator (Spectra-Physics Tsunami) is
the natural class, and its datasheet shows which line binds. At the
short end of its pulse-width range, a 2~ps sech$^{2}$ pulse
(time--bandwidth product 0.315) has an intensity-spectrum FWHM of
0.158~THz, i.e.\ $\sigma/2\pi = 0.095$~THz via Eq.~(\ref{eq:fwhm})
with the sech$^{2}$ spectrum treated as an effective Gaussian,
below the bandwidth window: at the 10~mm reference crystal this alone
gives $P = 0.66$. Bandwidth is the binding line---and a benign one,
because the purity is recovered by re-matching the crystal to
$L \approx 17$~mm rather than by modifying the laser. Every other
line passes trivially: the specified 0.2\%~rms intensity noise gives
a double-pair boost excess of only $4\times 10^{-6}$, and the 80~MHz
repetition-rate comb has teeth far too finely spaced to be resolved
by the JSA.

\subsection*{Consistency with published measurements}

The physical premises of the budgets are borne out by published
experiments, beginning with the multimode channel. Kwon
\emph{et al.}~\cite{kwon2009} pumped a 3~mm type-I BBO crystal with a
cw multimode 405~nm diode laser ($\sim$24 longitudinal modes within a
0.67~nm FWHM) and found that the pump coherence structure is
completely transferred to the biphoton interference: the $N = 2$
de Broglie (NOON) fringes show periodic wave-packet revivals at the
pump mode-spacing period with $\sim$98\% visibility. Their
signal--idler Hong--Ou--Mandel dip, by contrast, has visibility above
98\% with no revival---pair-exchange interference is insensitive to
the multimode structure. This is exactly the premise of our comb
channel: the pump modes imprint on the two-photon state along the
sum-frequency coordinate, where they leave signal--idler interference
untouched but directly damage the quantity our budgets bound, the
heralded single-photon purity probed by \emph{two-source}
interference. Since Ref.~\cite{kwon2009} does not measure heralded
purity, the comparison is premise-level rather than number-level.

For the observable itself, the flagship measurement is that of Mosley
\emph{et al.}~\cite{mosley2008}: two independent
group-velocity-matched KDP sources, pumped by femtosecond pulses at
415~nm and used without narrowband spectral filters, gave a four-fold
Hong--Ou--Mandel visibility of $94.4 \pm 1.6\%$ between independently
heralded photons---a raw value with no background subtraction,
presented as a lower bound on the mean photon purity, at heralding
efficiencies up to 44\%. This is precisely the quantity our budgets
constrain, since $V_{\mathrm{HOM}} = P$ for independent identical
sources. A singular-value decomposition of the measured joint
spectral intensity, under an assumed flat phase, instead gives
$\sim$98\%: even in a flagship source, real-world pump and phase
imperfections cost several percent---exactly the 1--5\% scale that
the budget rows of Table~\ref{tab:budget} translate into datasheet
tolerances.

Zielnicki \emph{et al.}~\cite{zielnicki2018} make the same point from
the characterization side: for one and the same unfiltered BBO
source, the inferred purity is technique-dependent, $0.88 \pm 0.02$
(Fourier-transform JSI) and $0.87 \pm 0.03$ (dispersive-fiber JSI)
against $0.66 \pm 0.02$ (from $g^{(2)}$) and $0.61 \pm 0.05$
(two-source HOM). JSI-based methods are blind to the joint spectral
phase---pump chirp in particular---and therefore overestimate the
purity, which empirically motivates the stated scope of our model:
a flat spectral phase throughout, with a chirp channel as a natural
future extension~\cite{sanchezlozano2012}. We note honestly what
these comparisons do not provide: none of these experiments publishes
purity as a function of a datasheet specification---precisely the gap
the budget tables fill---so the published record validates the
premises and scales of the model, not the budget numbers themselves.

\subsection*{Extensions and limitations}

Our ideal purity of 0.810 is the familiar ceiling of sinc-shaped
phase matching, set by the side lobes of the unapodized crystal.
Domain engineering and apodization reshape the phase-matching function
toward a Gaussian and raise this ceiling well above
0.81~\cite{branczyk2011,graffitti2018}. Our budgets are
\emph{relative} degradations from whatever ceiling the phase matching
provides, and the Monte-Carlo ensemble method applies unchanged to
engineered crystals: only the sinc factor in Eq.~(\ref{eq:jsa}) need
be replaced. This is now demonstrated rather than asserted: repeating
the full jitter sweep with the amplitude-matched Gaussian phase
matching of Eq.~(\ref{eq:phig}), whose ideal purity is 0.9818, yields
relative budgets of $30 \pm 0.3$/$68.5 \pm 0.8$/$102 \pm 1$~GHz~rms
(Table~\ref{tab:budget}), about 8\% \emph{larger} than the sinc
values at the 1\% level (8--13\% across the three levels), and these
are reproduced exactly, within one standard error, by the closed form
of Eq.~(\ref{eq:pj}), which predicts 29.9/69.0/101.7~GHz
(Fig.~\ref{fig:validation}). The sinc-model budgets quoted above are
therefore mildly conservative for engineered crystals.

The budgets also translate into two-photon interference, but the
translation depends on the architecture. For two \emph{independent}
identical heralded sources the Hong--Ou--Mandel visibility equals the
single-source purity, $V_{\mathrm{HOM}} = P$
\cite{hong1987,mosley2008}, so the jitter budgets above bound the
interference of sources pumped by separate free-running lasers. If
instead both sources are pumped by the \emph{same} laser, the drift
is common-mode: in each trial the two JSAs are translated rigidly
together by the same center shift, every realization retains the
ideal single-source purity, and within this model the shared-drift
HOM visibility stays at the ideal value $P_0$ \emph{exactly},
independent of the jitter amplitude. The reported budgets are
therefore conservative for single-laser multiplexed architectures, in
which center jitter cancels in common mode.

Several limitations bound the present results, each pointing to future
work. The model uses first-order GVM only: $\Delta k$ is linearized,
with no Sellmeier dispersion or higher-order terms, so the
phase-matching ridge is perfectly straight; curvature will matter for
broadband pumps and long crystals. We compute heralded purity only;
heralding efficiency and absolute pair rate are outside the model,
though a complete source assessment must weigh all three
jointly~\cite{meyerscott2020}. The quasi-static treatment of jitter
is no longer an open caveat: its regime of validity is bounded
explicitly in Sec.~\ref{sec:model}---noise slower than the
$\sim$1~ps pump coherence time and faster than the integration
time---and only sub-picosecond noise, which acts coherently within
each pulse, falls outside it on the fast side; drift slower than the
integration time is instead summarized deterministically through the
bounded-excursion conversion of Eq.~(\ref{eq:pp}). An incoherent pump linewidth,
beyond the Gaussian coherent bandwidth, is not yet a separate
mixed-state channel, which is why the Lorentzian linewidth in
Table~\ref{tab:datasheet} enters only as a quasi-static bound.
Similarly, the comb model asserts mutually incoherent lines and
therefore covers free-running multimode diodes only; a mode-locked
pump comb is phase-coherent, so its JSA is a coherent sum---a pure
state---and requires a coherent-sum treatment rather than the mixture
of Eq.~(\ref{eq:rho}). Finally, the treatment is collinear and
single-spatial-mode.

\section{Conclusion}

We have converted pump-laser datasheet specifications into a
quantitative imperfection budget for the heralded spectral purity of
an SPDC source, using a first-order GVM model with Monte-Carlo pump
ensembles validated against a closed-form double-Gaussian benchmark
that reproduces the measured Gaussian-phase-matching budgets exactly:
$30 \pm 0.3$/$68.5 \pm 0.8$/$102 \pm 1$~GHz measured versus
29.9/69.0/101.7~GHz predicted. For the reference source (10~mm
crystal, ideal purity 0.810), keeping the purity within 1\%/5\%/10\%
of ideal requires center-frequency jitter below
$27.7 \pm 0.3$/$62.7 \pm 0.7$/$91 \pm 1$~GHz~rms, and the budgets
obey a demonstrated $1/L$ law over 5--20~mm crystals (ratios
$1:2.00:4.00$), so halving the crystal doubles every budget at the
cost of pair rate. Single-longitudinal-mode operation is mandatory: a
single additional equal-power mode already costs 17\% of the purity.
Intensity noise, by contrast, leaves the purity exactly invariant and
instead inflates the double-pair boost. The rule for choosing a pump
is simple: spectral specifications damage purity, statistical
specifications damage pair statistics, and both columns of the
datasheet must be checked.

\begin{acknowledgments}
We thank Prof.\ Yas Al-Hadithi (King Abdulaziz University) for his
guidance and for discussions on laser physics.
\end{acknowledgments}
